---
title: "Data preparation"
author: "by: Frida van Teijlingen"
bibliography: references.bib
---

# Set up

```{r}
rm(list = ls())

#---- Loading packages ----
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

#---- Data saving ----
fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

#---- Loading in data ----
fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

#---- Output ----
fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

#---- Network data function ----
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}


#---- Gender scraper -----
gender_scraper.NV <- function(names = "names element", web_page = "https://nvb.meertens.knaw.nl/naam/is/") {

    names$first_name <- sapply(strsplit(names$Naam, " "), `[`, 1)  # This code should work as a way to extract first names from the ego characteristics dataset
    names$male <- NA
    names$female <- NA

    for (i in 1:nrow(names)) {

        # print(names$first_name[i])

        web_page <- read_html(paste0("https://nvb.meertens.knaw.nl/naam/is/", names$first_name[i]))

        table <- web_page %>%
            rvest::html_elements("body") %>%
            rvest::html_elements("table") %>%
            rvest::html_table()

        if (length(table) == 0) {

            print(length(table))

            names$male[i] <- NA
            names$female[i] <- NA

        } else {

            # print(table) print(table[[1]][[2,3]]) # Check if values for male are coherent and
            # accurate print(table[[1]][[6,3]]) # Check if values for female are coherent and
            # accurate

            names$male[i] <- as.numeric(ifelse(table[[1]][[2, 3]] == "--", 0, table[[1]][[2, 3]]))  # Make sure non-occurences are not registered as '--'
            names$female[i] <- as.numeric(ifelse(table[[1]][[6, 3]] == "--", 0, table[[1]][[6, 3]]))  # Make sure non-occurences are not registered as '--'

        }

    }  # end forloop

    names <- names %>%
        mutate(perc_female = case_when(is.na(female == TRUE) & is.na(male) == TRUE ~ NA, is.na(female) ==
            TRUE ~ 0, is.na(male == TRUE) ~ 1, .default = round((female/(male + female)), 2))) %>%
        select(!c(male, female, first_name))

    return(names)


}  # end function


```

# Libraries

```{r}
packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite")
fpackage.check(packages)
```

```{r}
# Use multiple functions across libraries
# igraph::triad.census()
# sna::triad.census()
```

# Loading data

```{r}
scholars <- fload("./data/scholars_20240924.rda")
```

# Data selection

## Select sociology departments of Dutch universities: df_allunis

```{r}
unidata <- fcolnet(data = scholars, 
                university = c("UU", "RUG", "UvA", "RU"),
                discipline = c("sociology"), 
                waves = list(c(2015, 2019), c(2020, 2024)), 
                type = c("first"))
```

## Creating df ego characteristics

```{r}
df_ego <- do.call(rbind.data.frame, unidata$data$demographics)
```

## Creating df works characteristics

```{r}
df_works <- do.call(rbind.data.frame, unidata$data$works)
```

# Data preparation

Adding and manipulating variables \## First names for Gender Scraper

```{r}
x <- data.frame(Naam = df_ego$Naam)

first_name <- sapply(strsplit(x$Naam, " "), `[`, 1)  # This code should work as a way to extract first names from the ego characteristics dataset



df_names <- data.frame(x, first_name, male = NA, female = NA)  # seem to have worked

head(df_names)

# df_names <- df_names[1:5,] # take first elements to try out

#test
i <- 1
df_names$first_name[i]
```

## Adding gender to ego characteristics

```{r}
# Function at top of document
df_ego <- gender_scraper.NV(names = df_ego, web_page = "https://nvb.meertens.knaw.nl/naam/is/")
```

## Gender distribution plotcheck

```{r}
gendercheck <- df_ego %>%
    count(perc_female)

plot(gendercheck$perc_female, gendercheck$n)
```

# Network visuals

## Wave 1: 2015-2019

```{r}
#wave 1
unis_w1 <- igraph::graph_from_adjacency_matrix(
  unidata$nets[1,,], #the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(unis_w1,
  vertex.label = NA,
  vertex.size = 3,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```

## Wave 2: 2020-2024

```{r}
#wave 2
unis_w2 <- igraph::graph_from_adjacency_matrix(
  unidata$nets[2,,], #the second wave of data.
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(unis_w2,
  vertex.label = NA,
  vertex.size = 3,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```

# Linking travel distance to universities

```{r}
traveltimetbl <- read_csv2("traveltime_ov.csv")

```

# Saving the prepared data

```{r}
#fsave(df_ego, file = "df_ego", location = "data/processed/")
#fsave(df_works, file = "df_works", location = "data/processed/")
#fsave(unidata, file = "unidata", location = "data/processed/")


```

# Code for loading in the prepared data

travelt_matrix in w6traveltimes.html

```{r}

#df_works <- fload("data/processed/df_works_20241009.rda")

#df_ego <- fload("data/processed/df_ego_20241009.rda")

#unidata <- fload("data/processed/unidata_20241009.rda")

#travelt_matrix <- fload("data/processed/travelt_matrix20241009")
```
