---
title: "Data & methods"
author: "by: Frida van Teijlingen"
bibliography: references.bib
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

# Set up
```{r}
rm(list = ls())
```

```{r warning=FALSE, globalsettings, echo=FALSE, results='hide'}

library(knitr)
library(formatR)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()


```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

***‘Out of sight, out of mind: a social network approach to geographical distance and co-authorship collaboration at four Sociology departments in the Netherlands’***

--------------------------------


## Functions
```{r}
#---- Loading packages ----
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

#---- Data saving ----
fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

#---- Loading in data ----
fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

#---- Output ----
fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}


#---- Moran's I ----
fMoran.I <- function(x, weight, scaled = FALSE, na.rm = FALSE, alternative = "two.sided", rowstandardize = TRUE) {
    if (rowstandardize) {
        if (dim(weight)[1] != dim(weight)[2])
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n)
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        ROWSUM <- rowSums(weight)
        ROWSUM[ROWSUM == 0] <- 1
        weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n *
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided")
            pv <- if (obs <= ei)
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater")
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    } else {
        if (dim(weight)[1] != dim(weight)[2])
            stop("'weight' must be a square matrix")
        n <- length(x)
        if (dim(weight)[1] != n)
            stop("'weight' must have as many rows as observations in 'x'")
        ei <- -1/(n - 1)
        nas <- is.na(x)
        if (any(nas)) {
            if (na.rm) {
                x <- x[!nas]
                n <- length(x)
                weight <- weight[!nas, !nas]
            } else {
                warning("'x' has missing values: maybe you wanted to set na.rm = TRUE?")
                return(list(observed = NA, expected = ei, sd = NA, p.value = NA))
            }
        }
        # ROWSUM <- rowSums(weight) ROWSUM[ROWSUM == 0] <- 1 weight <- weight/ROWSUM
        s <- sum(weight)
        m <- mean(x)
        y <- x - m
        cv <- sum(weight * y %o% y)
        v <- sum(y^2)
        obs <- (n/s) * (cv/v)
        if (scaled) {
            i.max <- (n/s) * (sd(rowSums(weight) * y)/sqrt(v/(n - 1)))
            obs <- obs/i.max
        }
        S1 <- 0.5 * sum((weight + t(weight))^2)
        S2 <- sum((apply(weight, 1, sum) + apply(weight, 2, sum))^2)
        s.sq <- s^2
        k <- (sum(y^4)/n)/(v/n)^2
        sdi <- sqrt((n * ((n^2 - 3 * n + 3) * S1 - n * S2 + 3 * s.sq) - k * (n * (n - 1) * S1 - 2 * n *
            S2 + 6 * s.sq))/((n - 1) * (n - 2) * (n - 3) * s.sq) - 1/((n - 1)^2))
        alternative <- match.arg(alternative, c("two.sided", "less", "greater"))
        pv <- pnorm(obs, mean = ei, sd = sdi)
        if (alternative == "two.sided")
            pv <- if (obs <= ei)
                2 * pv else 2 * (1 - pv)
        if (alternative == "greater")
            pv <- 1 - pv
        list(observed = obs, expected = ei, sd = sdi, p.value = pv)
    }


}


```

## Libraries

```{r, warning=FALSE, message=FALSE, results='hide'}
packages <- c("RSiena", "tidyverse", "stringdist", "stringi", "igraph", "leaflet")
fpackage.check(packages)
```

## Load in data

```{r}
#df_ego
df_ego <- fload("./data/processed/df_ego_20241009.rda")

#df_works
df_works <- fload("data/processed/df_works_20241009.rda")

#df_allunis
unidata <- fload("data/processed/unidata_20241009.rda")

travelt_matrix <- fload("data/processed/travelt_matrix_20241009.rda")
```

-------------------------------

# Data

Ahead of answering the research questions laid a process of cleaning and transforming raw data into data that was ready to be analysed. The dataset provided as part of the course material consisted of scholars within the Sociology and Political Science departments of Dutch universities. This data was previously webscraped from OpenAlex, an open access bibliographic catalogue of scientific papers, authors and institutions. The scraped data includes names, university of employment, job description, Google scholar ID, discipline and OpenAlex ID. This data was available for both 2022 and 2024, as this provides insight into affiliations of scholars working at certain universities at both time points. I made a selection of four sociology departments of Dutch universities, namely Utrecht University (UU), Radboud University (RU), University of Amsterdam (UvA), and the University of Groningen (RUG). This selection was based on their partship of the ICS (Interuniversity Centre for Social Science Theory and Methodology), a cooperation initiative for social science research. Due to this shared membership, it would be likely that these four universities collaborate often, making it an interesting selection to study whether geographical distance influences these collaborations. In order to analyse changes within the networks of scholars and to answer my research questions, I made a distinction between two publication waves, ranging from 2015 to 2019 (wave 1) and 2020 to 2024 (wave 2). The publications within these waves were the basis for the collaboration networks.

The selection of the two waves of the four universities resulted in a total sample of 234 scholars (UU, N= 48; RU, N=44; UvA, N=59; RUG, N=67; 16 scholars were not apart of any of these departments at the 2022 measure point). The 'Data preparation - code' page in this subtab provides the code used for the preparation of the data thus far.

```{r}
df_ego %>%
mutate(University = case_when(str_detect(Universiteit1.22, "RUG") ~ "RUG",
                              str_detect(Universiteit1.22, "RU") ~ "RU", 
                              str_detect(Universiteit1.22, "UU") ~ "UU", 
                              str_detect(Universiteit1.22, "UvA") ~ "UvA", 
                              .default = "Not in Department")) %>%
    count(University) %>%
    fshowdf()

```

## Independent variable

The independent variable in this study is geographical distance, which will be treated as a time constant variable. I used travel time by public transport as a proxy measurement for the geographical distance between the universities, as I believe this to be a sufficient measurement due to its similar ‘psychic distance’: a great geographical distance and long travel time will likely be experienced as the same felt distance. Additionally, Dutch universities are generally very accessible by trains and busses and travelling by public transport allows for working during one’s commute, which is why I think a majority of scholars will opt to travel by public transport over other forms of transportation. The travel time in minutes was gathered by estimating the travel time via the Dutch Railways travel planning website (<https://www.ns.nl/reisplanner/>) between each of the universities, departing at 8AM on the 7th of October 2024. A looping code was used to match the travel time between the universities to each scholar and their university of employment. The 'Matching and looping travel times - code' page in this subtab provides the code used for the preparation and creation of this variable.
The map below gives an overview of the geographical locations of the universities that are included in this study, as well as the travel times by use of public transport between each university.


```{r}
# Define university coordinates
uni_coords <- data.frame(
  uni_label = c("UU", "RUG", "UvA", "RU"),
  lng = c(5.171430, 6.557490, 4.912590, 5.8623425),
  lat = c(52.084460, 53.222250, 52.363210, 51.8192069)
)

# Define travel times between universities
# Each row represents "from", each column represents "to"
map_tt <- data.frame(
  from = c("UU", "UU", "UU", "RUG", "RUG", "UvA"),
  to = c("RUG", "UvA", "RU", "UvA", "RU", "RU"),
  time = c(156, 63, 91, 153, 180, 109)
)
```


```{r}
map <- leaflet() %>%
  addProviderTiles(provider = "Stadia") %>%
  setView(lng = 5.2913, lat = 52.1326, zoom = 7.4) %>%
  addMarkers(
    lng = uni_coords$lng,
    lat = uni_coords$lat,
    label = uni_coords$uni_label
  )

#Loop through travel times to draw lines and add labels
for (i in 1:nrow(map_tt)) {
  from_coords <- uni_coords[uni_coords$uni_label == map_tt$from[i], c("lng", "lat")]
  to_coords <- uni_coords[uni_coords$uni_label == map_tt$to[i], c("lng", "lat")]  #Get coordinates of "from" and "to" universities
  
#Draw a straight line between them
  map <- map %>%
    addPolylines(
      lng = c(from_coords$lng, to_coords$lng),
      lat = c(from_coords$lat, to_coords$lat),
      color = "blue",
      weight = 2,
      opacity = 0.3
    )
  
#Calculate midpoint for placing the travel time label
  mid_lng <- mean(c(from_coords$lng, to_coords$lng))
  mid_lat <- mean(c(from_coords$lat, to_coords$lat))
  
#Add a label showing travel time at the midpoint of the line
  map <- map %>%
    addLabelOnlyMarkers(
      lng = mid_lng,
      lat = mid_lat,
      label = paste(map_tt$time[i], "min"),
      labelOptions = labelOptions(noHide = TRUE, direction = "top", textOnly = TRUE)
    )
}

map

```






## Dependent variables

1.  Co-authorship / publications 

The main dependent variable in this study is co-authorship of a published paper. A tie between scholars is assumed when they co-authored a paper together. I decided to include directed ties from the first author, meaning that the first author listed on the publication is assumed to be the one sending the invitation for collaboration to the other author(s) on the publication. I made this decision based on the assumption that first authors are likely to take the lead in the collaboration process, and thus the one that will generally be expected to visit co-authors during the collaboration more often than the co-authors, thus increasing the relative importance of geographical distance for the likelihood of tie-formation.



2.  Gender 

Gender is not part of the Open Alex data and thus not part of the initial dataset. Therefore, gender was gathered by the use of webscraping, a technique to extract information from publicly available webpages. In this study, I made use of the Gender Scraper developed by Niels Vullings. In short, this webscrape function gave a score ranging between ‘0’ and ‘1’ to indicate the proportion of women in the Netherlands with the same first name, ‘0’ meaning there were no women with the same name, ‘1’ meaning that all people with the same name are women.

```{r}
summary(df_ego$perc_female)
```


## Ethical considerations

The use of webscraping as a form of data collection often raise some ethical concern regarding privacy and consent. While a lot of data, like the data used here, is publicly available, many individuals might not fully consent to their data being collected and analysed for research purposes and webscraping thereby violates their expected privacy. As researchers, we are therefore obligated to conform to terms of service set on the platform of which the data was obtained, anonymize the data, and to be thorough and considerate of which data is ethically acceptable to webscrape.

------------------------------------------------------------------------

# Methods

To test the before mentioned hypotheses and thereby answering the research questions of this study, this study takes a structural approach. First, a descriptive analysis is done to identify the structure of the networks. In this first stage, statistics related to network size, degrees, transitivity, dyad and triad census, and Moran’s I are analysed. The results of this analyses can be found below. Next, an exploratory analysis was conducted using RSiena. RSiena (Simulation Investigation for Empirical Network Analysis) is a computer program fit for R that performs statistical estimations of the evolution of social network models, thus suitable for this study (Ripley et al., 2011). 

A step wise approach was used to build the models. The null model will solely contain the reciprocity and density of the network. In Model 1 and 2, covariate effects are added. Model 1 will add dyadic covariate effects of university affiliation, in Model 2 the actor-level covariate effects and dyadic covariate effects of gender will be added. In these models, the structural effects are ignored in order to solely see the ego-characteristic related effects. The structural effects will be included in Model 3 to give insight into behavioural patterns within the network.



# Descriptive statistics

## Network visualization

### Wave 1: 2015-2019 vs Wave 2: 2020-2024

The two graphs below visualize the collaboration networks wihtin the two waves. It includes all scholars at Sociology departments at University Utrecht (yellow), Radboud University (red), University of Amsterdam (darkgreen), and Rijksuniversiteit Groningen/RUG (blue). From these graphs it becomes apparent that in both waves, there are obvious clusters of collaboration within universities. Additionally, these graphs suggest that there is also some form of clustering between certain universities, where scholars from the same university show a preference in the universities of scholer with whom they tend to collaborate.

```{r}

kleurtjes <- ifelse(df_ego$Universiteit1.22 == "UU", "yellow", "blue" )
kleurtjes <- ifelse(df_ego$Universiteit1.22 == "RU", "red", kleurtjes )
kleurtjes <- ifelse(df_ego$Universiteit1.22 == "UvA", "darkgreen", kleurtjes )

#wave 1
unis_w1 <- igraph::graph_from_adjacency_matrix(
  unidata$nets[1,,], #the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

#wave 2
unis_w2 <- igraph::graph_from_adjacency_matrix(
  unidata$nets[2,,], #the second wave of data.
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.rownames = NULL
)
 par(mfrow = c(1, 2))
 
#wave 1
plot(unis_w1,
  vertex.label = NA,
  vertex.color = kleurtjes,
  vertex.size = sqrt(igraph::degree(unis_w1))*5,
  edge.width = 0.2,
  edge.arrow.size = 0.2,
  main = "Collaboration network - Wave 1")

#wave 2
plot(unis_w2,
  vertex.label = NA,
  vertex.color = kleurtjes,
  vertex.size = sqrt(igraph::degree(unis_w2))*5,
  edge.width = 0.2,
  edge.arrow.size = 0.2,
  main = "Collaboration network - Wave 2")


par(mfrow = c(1, 1))
```

### Wave 1: 2015-2019 vs Wave 2: 2020-2024 - Without isolates

The two graphs below are similar, but the actors that were isolate (meaning no ties) in both waves were excluded. Between wave 1 and 2, many actors that were isolates during wave 1 have since formed ties during wave 2.

```{r}
 par(mfrow = c(1, 2))

# Identify isolates in waves 1 and 2
isolates_w1 <- igraph::V(unis_w1)[igraph::degree(unis_w1) == 0]
isolates_w2 <- igraph::V(unis_w2)[igraph::degree(unis_w2) == 0]

# Find common isolates in both waves
common_isolates <- intersect(isolates_w1, isolates_w2)

# Remove common isolates from waves 1 and 2
unis_w1_clean <- igraph::delete.vertices(unis_w1, common_isolates)
unis_w2_clean <- igraph::delete.vertices(unis_w2, common_isolates)
 
#wave 1
plot(unis_w1_clean,
  vertex.label = NA,
  vertex.color = kleurtjes,
  vertex.size = sqrt(igraph::degree(unis_w1))*5,
  edge.width = 0.2,
  edge.arrow.size =0.2,
  main = "Collaboration network - Wave 1",
  sub = "no isolates")

#wave 2
plot(unis_w2_clean,
  vertex.label = NA,
  vertex.color = kleurtjes,
  vertex.size = sqrt(igraph::degree(unis_w2))*5,
  edge.width = 0.2,
  edge.arrow.size =0.2,
  main = "Collaboration network - Wave 2",
  sub = "no isolates")



par(mfrow = c(1, 1))

```

## Size

For the next descriptive analyses, we will look at the full network, including the isolates. There are a total of 234 scholars in this network, both in wave 1 and 2. The network in wave 1 contains 145 edges, reflecting the number of connections between scholars. Wave 2 contains 258 edges, indicating a growth in the density of the network compared to wave 1.

```{r}
# number of nodes
igraph::vcount(unis_w1)
# number of edges
igraph::ecount(unis_w1)

# number of nodes
igraph::vcount(unis_w2)
# number of edges
igraph::ecount(unis_w2)
```

## Degrees

The histograms show the distribution of in-degrees (incoming edges) in the network of both waves. Degrees can give insight into how many scholars have few or many incoming connections, which can foreshadow some aspects of centralization or popularity within the networks.

```{r}
#igraph::degree(unis_w1)
hist(table(igraph::degree(unis_w1)), xlab='indegree', main= 'Histogram of indegree wave 1')

#igraph::degree(unis_w2)
hist(table(igraph::degree(unis_w2)), xlab='indegree', main= 'Histogram of indegree wave 2')
```

## Transitivity

Transitivity is a measure for the extent to which the actors in the network tend to cluster together. In the context of this study, it assesses the likelihood that two connected scholars have a mutual connection to a third scholar. The probability of a mutual connection to a third scholar is 0,18 in wave 1 and a little bit lower at 0,17 in wave 2.

```{r}
igraph::transitivity(unis_w1)
igraph::transitivity(unis_w2)
```

## Dyad census

Wave 1 contains a total of 19 reciprocal (two-way) ties, wave 2 a total of 30. Looking at the assymetic ties, there is a noticable increase from 107 in wave 1 to 198 in wave 2.

```{r}
igraph::dyad_census(unis_w1)
igraph::dyad_census(unis_w2)
```

## Triad census

Comparing wave 1 and 2, we can see a few interesting changes: 1. The occurrence of a a cycle of three nodes where two are reciprocally connected and one node connects outward to a third is more than doubled in wave 2.\
2. While for almost all triad configurations the occurrence increase in wave 2, there is one configuration that decreases from 3 to 1: configuration 120C, a cyclic triad where one of the ties is reciprocal, the other two are single directed ties. 3. The number of complete triads (triad configuration 300, meaning that all nodes are reciprocally connected, each node has ties to the other two in both directions) increased from 1 to 2.

```{r}
igraph::triad_census(unis_w1)
```

```{r}
igraph::triad_census(unis_w2)
```

## Moran’s I

Moran's I is a measurement that is used to determine whether nodes who are closer to one another in the network are also more alike. In the context of this study, I analyse whether scholars that are employed at the same university, also share a similar pattern in collaboration with scholars from other universities in terms of geographical proximity. For example, do scholars that are all employed at Radboud University share a similar collaboration pattern, such as collaborating relatively more with scholars from Utrecht University compared to the other universities? Or in other words: Is the network segregated based on collaboration patterns?

I opted to not rowstandardize in order to capture the absolute rather than relative influence. This reflects the global structure of the network more directly, and it allows for an increase in influence when there are more connections to similar others compared to when there are just a couple of connections to similar others.

The observed Moran's I of wave 1 shows a positive and significant value of 0.32, indicating a moderate degree of clustering of similar collaboration densities for scholars and similar others. The observed Moran's I of wave 2 is also a positive and significant value of 0.82, indicating high spatial autocorrelation.

The first hypothesis of this study suggests that geographical proximity influences collaboration patterns of scholars at the Sociology departments at four Dutch universities. The positive and significant Moran's I statistics indicate that scholars show a tendency to form denser collaborative ties to scholars at universities that are spatially closer compared to scholars at universities that are spatially further away. The results of these two Moran's I statistics provide corroborating evidence for the first hypothesis. 

```{r}
#wave 1
require(sna)
geodistances <- geodist(unidata$nets[1,,], count.paths = TRUE)
geodistances <- geodistances$gdist

# first define a nb based on distance 1.
weights1 <- geodistances == 1

MIw1 <- fMoran.I(travelt_matrix [, 1], scaled = FALSE, weight = weights1, na.rm = TRUE, rowstandardize = FALSE)

MIw1
```

```{r}
#wave 2
require(sna)
geodistances <- geodist(unidata$nets[2,,], count.paths = TRUE)
geodistances <- geodistances$gdist

# first define a nb based on distance 1.
weights1 <- geodistances == 1

MIw2 <- fMoran.I(travelt_matrix [, 2], scaled = FALSE, weight = weights1, na.rm = TRUE, rowstandardize = FALSE)

MIw2
```

Hypothesis 2 suggested that the influence of geographical distance on between university collaboration of Dutch Sociology departments has increased over time. By comparing the above calculated Moran's I statistics, we can test whether the Moran's I has increased significantly over time between wave 1 and 2. Based on the test statistics below, it can be concluded that there is a high and significant difference between the two waves. The increase in the Moran's I suggests that the network autocorrelation has increased over time. This implies that the network has become more clustered and similar, which indicates less division based on geographical distance. Hypothesis 2 is thereby also accepted.

```{r}
psych::paired.r(xy= MIw1$observed, xz= MIw2$observed, n = 100)
```

# References
Ripley, R. M., Snijders, T. A., Boda, Z., Vörös, A., & Preciado, P. (2011). Manual for RSIENA. University of Oxford, Department of Statistics, Nuffield College, 1, 2011.