---
title: "Introduction"
author: "by: Frida van Teijlingen"
bibliography: references.bib
---

```{=html}
<style>
body {
text-align: justify}
</style>
```
```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}

library(knitr)
library(formatR)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

***‘Out of sight, out of mind: a social network approach to geographical distance and co-authorship collaboration at four Sociology departments in the Netherlands’***


# Introduction

Over the last few centuries, globalization processes have resulted in greater economic, political and cultural connections and interdependence. According to some, the globalization of science is one of the most important, if not the most important, aspect of modern globalization (Freeman, 2010). With a globalizing world also comes an immense (and growing) worldwide network of international connections and collaboration. The subsequent knowledge flows as a result of these connections and collaborations are experienced as beneficial for economic development, specialization, networking, and the competitiveness in the scientific field (Gui et al., 2019). Additionally, on an smaller level, the large international collaboration network can benefit both students and teachers when seeking specific knowledge and specializations (Fitzgerald et al., 2021).

When seeking collaboration, geographical distance can be experienced as a barrier. A greater distance can affect the effectiveness and efficiency of collaboration, if it does not prevent collaboration entirely in the first place (Pallot et al., 2010). While globalization is not able to shrink geographical distances in a literal sense, our individual perception and awareness of the distance between us and a secondary location, also referred to as ‘psychic distance’, has decreased as a result of globalization (Ojala, 2009). That is, phones, e-mail, and other telecommunication technologies have made it much easier to reach out to others who are not in our immediate vicinity, advancing collaboration opportunities (Hoekman et al., 2010). While this relation has been proven for long-distance, international scientific collaboration (e.g. Hoekman et al., 2010), little research has been done on the relationship between geographical distance and scientific collaboration networks on a single country level. While it can be expected to be less evident, geographical distance can still be expected to be an affecting factor in scientific collaboration networks between actors within a country, as travel time and convenience can still be of relevant influence. Studying this on a single country level can give us more insight into the degree of influence geographical distance can have on scientific collaboration.

This study aims to do so by studying the influence of geographical distance on co-authorship collaboration networks in the Netherlands, specifically between the sociology departments at several Dutch universities. This leads to the following research question: *To what extent does the geographical distance between Sociology departments at Dutch universities influence the between-university collaboration network of scientists located at these universities* [RQ1]*?*

Broad societal changes, such as the COVID-19 pandemic between 2020 and 2022, likely had lasting consequences for scientific collaboration networks. That is, government mandated measures such as working from home, limited travelling and avoiding close contact to others, might have changed the way these networks were constructed. This study will therefore also look at changing spatial patterns of research collaboration by studying co-authorship over time, resulting in the following questions: *To what extent has the relation between geographical distance and between-university collaboration networks of Dutch Sociology departments changed over time* [RQ2]*?* And: *What patterns can be discovered within the collaboration network and within the changes over time* [RQ3]*?*

## Introduction - References

Fitzgerald, J., Ojanperä, S., & O’Clery, N. (2021). Is academia becoming more localised? The growth of regional knowledge networks within international research collaboration. Applied Network Science, 6(1), 1-27. <https://doi.org/10.1007/s41109-021-00371-w> Freeman, R. B. (2010). Globalization of scientific and engineering talent: international mobility of students, workers, and ideas and the world economy. Economics of Innovation and New Technology, 19(5), 393-406. <https://doi.org/10.1080/10438590903432871> Gui, Q., Liu, C., & Du, D. (2019). Globalization of science and international scientific collaboration: A network perspective. Geoforum, 105, 1-12. <https://doi.org/10.1016/j.geoforum.2019.06.017> Hoekman, J., Frenken, K., & Tijssen, R. J. (2010). Research collaboration at a distance: Changing spatial patterns of scientific collaboration within Europe. Research policy, 39(5), 662-673. <https://doi.org/10.1016/j.respol.2010.01.012> Ojala, A. (2009). Is globalization reducing distances between countries? Some empirical evidence from foreign software firms operating in Japan. Journal of Japan Academy for Asian Market Economies, 12(1). Pallot, M., Bergmann, U., Kühnle, H., Pawar, K. S., & Riedel, J. C. (2010, June). Collaborative working environments: Distance factors affecting collaboration. In 2010 IEEE International Technology Management Conference (ICE) (pp. 1-10). IEEE. <https://doi.org/10.1109/ICE.2010.7476990>.
