---
title: "Week 2"
#bibliography: references.bib
author: "Your Name"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

## Research Questions
**Geographical distance and co-authorship collaboration network analysis**

Fitzgerald, J., Ojanperä, S., & O’Clery, N. (2021). Is academia becoming more localised? The growth of regional knowledge networks within international research collaboration. Applied Network Science, 6(1), 1-27. 
[Link to article Fitzgerald et al., 2021](https://www.springernature.com/gp/open-research/about/the-fundamentals-of-open-access-and-open-research)

--> "while communities present in the global collaboration network before 2000 were often based on historical geopolitical or colonial lines, in more recent years they increasingly align with a simple partition of countries by regions. These findings are unexpected in light of a presumed continual increase in globalisation, and have significant implications for the design of programmes aimed at promoting international research collaboration and knowledge diffusion."

Ponds, R., Van Oort, F., & Frenken, K. (2007). The geographical and institutional proximity of research collaboration. Papers in regional science, 86(3), 423-444.
[Link to article Ponds et al., 2007](https://www.sciencedirect.com/science/article/pii/S1056819023023989)

--> Dutch study. Studied research collaboration between universities, but also between universities and organizations. 
Geographical proximity important factor in research collaboration due to: trust, transfering tacit knowledge, deeper understanding, etc.


**RQ1: To what extent does the geographical distance between Dutch universities influence the between-university collaboration network of scientists in Sociology and Data Science departments within these universities?**:  

- Has this changed over time?  
- Does the influence of geographical distance differ between the two departments?  
- Does the influence of geographical distance affect scientists from various cohorts differently? But, age effect  
- Can a pattern be observed? Possibly affected by university rankings? Increase during/after the pandemic? Regional specialisation in related scientific subfields?    


**RQ2: To what extent does the influence of geographical distance differ between scientists who worked at more than one university compared to those who have worked at one?**:  

- Hmm  
- But, age effect  
- Are there even scientists who have only worked at one (except those just starting out) --> maybe study difference between >1 department vs just 1 department?  





Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----
  
This is just an example page of your `r colorize("lab journal", "red")`.
You can journal the assignments of the first day here.
  

----


```{r}
print("Hello Lab Journal")
```



