---
title: "Week 2"
#bibliography: references.bib
author: "Frida van Teijlingen"
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Research Questions

**Topic: Geographical distance and co-authorship collaboration network analysis**

Fitzgerald, J., Ojanperä, S., & O’Clery, N. (2021). Is academia becoming more localised? The growth of regional knowledge networks within international research collaboration. Applied Network Science, 6(1), 1-27. 
[Link to article Fitzgerald et al., 2021](https://www.springernature.com/gp/open-research/about/the-fundamentals-of-open-access-and-open-research)

--> "while communities present in the global collaboration network before 2000 were often based on historical geopolitical or colonial lines, in more recent years they increasingly align with a simple partition of countries by regions. These findings are unexpected in light of a presumed continual increase in globalisation, and have significant implications for the design of programmes aimed at promoting international research collaboration and knowledge diffusion."

Ponds, R., Van Oort, F., & Frenken, K. (2007). The geographical and institutional proximity of research collaboration. Papers in regional science, 86(3), 423-444.
[Link to article Ponds et al., 2007](https://www.sciencedirect.com/science/article/pii/S1056819023023989)

--> Dutch study. Studied research collaboration between universities, but also between universities and organizations. 
Geographical proximity important factor in research collaboration due to: trust, transfering tacit knowledge, deeper understanding, etc.


**RQ1: To what extent does the geographical distance between Dutch universities influence the between-university collaboration network of scientists in Sociology and Data Science departments within these universities?**:  

- Has this changed over time?  
- Does the influence of geographical distance differ between the two departments?  
- Does the influence of geographical distance affect scientists from various cohorts differently? But, age effect  
- Can a pattern be observed? Possibly affected by university rankings? Increase during/after the pandemic? Regional specialisation in related scientific subfields?    


**RQ2: To what extent does the influence of geographical distance differ between scientists who worked at more than one university compared to those who have worked at one?**:  

- Hmm  
- But, age effect  
- Are there even scientists who have only worked at one (except those just starting out) --> maybe study difference between >1 department vs just 1 department?  



----
  
**Introduction – Geographical distance and co-authorship collaboration network analysis**  
Over the last few centuries globalization processes have resulted in economic, political and cultural interdependence, supported by an immense worldwide network of international connections and collaboration (SOURCE). 
According to some, the globalization of science is the most important aspect of modern globalization (Freeman, 2010). The subsequent knowledge flows are beneficial for economic development and specialization, networking, and the competitiveness in the scientific field (Gui et al., 2019). Additionally, globalization has created a large international collaboration network that benefits both students and teachers when seeking specific knowledge and specializations (Fitzgerald et al., 2021)


While globalization might not have shrunk geographical distances in a literal sense, relative distance might actually have (SOURCE?). Intuitively, the relative distance to other individuals or organizations located in a different region, country, or even continent has become the shortest it has ever been. 
Communication methods, such as phones, e-mail, and other communication methods have made it much easier to reach out to others who are not directly in our immediate environment. 
--smaller: also within country!!
Combining the aspects of the globalization of science and a decrease of the relative distance between ourselves and others results in the main focus of this study, namely the extent to which  

Relevance of Social Networks perspective:
-	Use SNA to display the structure of the network




Fitzgerald, J., Ojanperä, S., & O’Clery, N. (2021). Is academia becoming more localised? The growth of regional knowledge networks within international research collaboration. Applied Network Science, 6(1), 1-27. https://doi.org/10.1007/s41109-021-00371-w 
Freeman, R. B. (2010). Globalization of scientific and engineering talent: international mobility of students, workers, and ideas and the world economy. Economics of Innovation and New Technology, 19(5), 393-406. https://doi.org/10.1080/10438590903432871 
Gui, Q., Liu, C., & Du, D. (2019). Globalization of science and international scientific collaboration: A network perspective. Geoforum, 105, 1-12. https://doi.org/10.1016/j.geoforum.2019.06.017 




Descriptive Q: What does the collaboration network of scientists of Sociology departments at Dutch universities look like?


**RQ1:**  
To what extent does the geographical distance between Sociology departments of Dutch universities influence the between-university collaboration network of scientists within these universities?

**RQ2: Case study type question**  
To what extent does the geographical distance within the Sociology department at * University influence the within-department collaboration network of scientists within this department?
*check for a university that is geographically spread

or: To what extent does the geographical distance between the Sociology department and Political science department at * University influence the between-department collaboration network of scientists within this department?






Sample: staff of sociology departments at Dutch universities

RQ1 Needs:  

-	Who published with who? # of publications for a weighted tie  
-	Where did they publish; e.g. prestige of the journals for additional RQ’s maybe  
-	Basic descriptives for possible control variables (age, sex, job title? Would be interesting)  
-	Distances in KM from one university to the other, via public roads vs hemelsbreed, why  
- ranking 'strength' of the collaboration: e.g. 1st/last vs else author position?  



If possible:  
-	Work or study at other universities  



RQ2 Needs:  

- Distances in KM from building to building  








----


# Assignments during class

## making a matrix and adding ties manually
```{r}
allzeromatrix <- matrix(0, nrow=4, ncol=4)

allzeromatrix

#anderemanier <- c(0, 0, 1, 0, 
                  #0, 0, 0, 0,
                  #0, 0, 0, 0,
                  #0, 0, 0, 0)
#anderemaniermatrix <- matrix(anderemanier, nrow=4, ncol=4)


#placing a 1 instead of 1 of the zeros
onetiematrix <- matrix(0, nrow=4, ncol=4)
onetiematrix[1, 3] <- 1 #from node 1 to 3
onetiematrix
```

## triad census, defining/coding the types of ties
```{r}
#install.packages("igraph")
library(igraph)

#triad_census(graph) #doesnt work yet, need to provide graph object
testgraph0<- graph_from_adjacency_matrix(adjmatrix = allzeromatrix)
triad_census(testgraph0) # gives 16 numbers, relate to 16 possible triads, e.g. 400000000000 means there are 4 003 triads aka no ties


testgraph1<- graph_from_adjacency_matrix(adjmatrix = onetiematrix)
triad_census(testgraph1)
```

## trying out descriptive statistics
```{r}
#install.packages("RSiena")
library(RSiena)
s501

dim(s501)

#selecting the first 10 by 10
trying1 <- s501[1:10, 1:10]

dim(trying1)

tryinggraph1<- graph_from_adjacency_matrix(adjmatrix = trying1)
triad_census(tryinggraph1)
plot(tryinggraph1)


# 82  0 37  0  0  0  0  0  0  0  0  0  0  0  0  1
#so, 82 non-ties, 37 reciprocated ties (of tryads! so dyad+one nonconnected counts), 1 complete circular tie
#total 120: (10*9*8)/2
```

```{r}
#centrality measures: degree, betweenness, closeness

#DEGREE
#table with degrees distribution
degree(tryinggraph1)
degree_distribution(tryinggraph1)

#normalized degree centrality:
#?hmmm...

#BETWEENNESS
betweenness(tryinggraph1)
betweenness.estimate(tryinggraph1, directed = TRUE, cutoff=10)
estimate_betweenness(tryinggraph1,  directed = TRUE, cutoff=10)


#CLOSENESS
closeness(tryinggraph1, mode = "out")

#mode: can switch between "out"degree, "in" degree, or "total"/"all"

#transitivity
transitivity(tryinggraph1)



```

```


----


Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>






