# Getting started
```{r}
rm(list = ls())


fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite")
fpackage.check(packages)
```

```{r}
scholars <- fload("./data/scholars_20240924.rda")
```

```{r}
# (sociology, RU)
demographics <- do.call(rbind.data.frame, scholars$demographics)
demographics <- demographics %>%
    mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
        is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
        ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
        is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

sample <- which((demographics$Universiteit1.22 == "RU" | demographics$Universiteit2.22 == "RU" | demographics$Universiteit1.24 ==
    "RU" | demographics$Universiteit2.24 == "RU") & (demographics$discipline.22 == "sociology" | demographics$discipline.24 ==
    "sociology"))

demographics_soc <- demographics[sample, ]
scholars_sel <- lapply(scholars, "[", sample)
```

```{r}
ids <- demographics_soc$au_id
wave2 <- wave1 <- matrix(0, nrow = length(ids), ncol = length(ids), dimnames = list(ids, ids))
```

```{r}
works <- scholars_sel$work

works_id <- unlist(lapply(works, function(l) l$id))
works_author <- unlist(lapply(works, function(l) l$author), recursive = FALSE)
works_year <- unlist(lapply(works, function(l) l$publication_year), recursive = FALSE)

df_works <- tibble(works_id, works_author, works_year)

dups <- which(duplicated(works_id))
# why are there some many papers of Batenburg duplicates

df_works <- df_works[-dups, ]
df_works_w2 <- df_works[df_works$works_year > 2019, ]
```

```{r}
# first to others
for (i in 1:nrow(df_works_w2)) {
    ego <- df_works_w2$works_author[i][[1]]$au_id[1]
    alters <- df_works_w2$works_author[i][[1]]$au_id[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
}

# last to others
for (i in 1:nrow(df_works_w2)) {
    ego <- rev(df_works_w2$works_author[i][[1]]$au_id)[1]
    alters <- rev(df_works_w2$works_author[i][[1]]$au_id)[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
}

# undirected
for (i in 1:nrow(df_works_w2)) {
    egos <- df_works_w2$works_author[i][[1]]$au_id

    if (sum(ids %in% egos) > 0) {
        wave2[which(ids %in% egos), which(ids %in% egos)] <- 1
    }
}
```

About the parameters:

    data: our scholars file
    university: Character vector with names of universities. We have several universities in the Netherlands. See above for relevant names.
    discipline: Character vector, either sociology or political science or both.
    waves: a list of numeric vectors with start and end year of wave.
    type:
        "first": directed: first author sending to others
        "last": directed: last author sending to others
        "all": undirected: ties between all authors

Output: - a list
- $nets: array of nomination networks.
- $data: sample of data (scholars)
```{r}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```



```{r}
#save the output of your function
test <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))




test_w1 <- igraph::graph_from_adjacency_matrix(
  test$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(test_w1,
  vertex.label = NA,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```
# Select sociology departments of Dutch universities: df_allunis
```{r}
df_allunis <- fcolnet(data = scholars, 
                university = c("UU", "RUG", "UvA", "EUR", "Leiden", "RU", "VU", "UvT", "TU Delft", "UvH"),
                discipline = c("sociology"), 
                waves = list(c(2018, 2019), c(2020, 2021), c(2022, 2023)), 
                type = c("first"))



```


```{r}
#wave 1
allunis_w1 <- igraph::graph_from_adjacency_matrix(
  df_allunis$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(allunis_w1,
  vertex.label = NA,
  vertex.size = 3,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```


```{r}
#wave 2
allunis_w2 <- igraph::graph_from_adjacency_matrix(
  df_allunis$nets[2,,], #for this example I take the second wave of data.
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(allunis_w2,
  vertex.label = NA,
  vertex.size = 3,
  edge.width = 0.2,
  edge.arrow.size =0.2)


```
```{r}
#wave 3
allunis_w3 <- igraph::graph_from_adjacency_matrix(
  df_allunis$nets[3,,], #for this example I take the second wave of data.
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(allunis_w3,
  vertex.label = NA,
  vertex.size = 3,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```


# webscraping gender
```{r}
df <- df_allunis$data
df_ego <- do.call(rbind.data.frame, df$demographics)
```

## first names
```{r}
x <- data.frame(Naam = df_ego$Naam)

first_name <- sapply(strsplit(x$Naam, " "), `[`, 1)  # This code should work as a way to extract first names from the ego characteristics dataset



df_names <- data.frame(x, first_name, male = NA, female = NA)  # seem to have worked

head(df_names)

# df_names <- df_names[1:5,] # take first elements to try out

#test
i <- 1
df_names$first_name[i]
```


```{r}
gender_scraper.NV <- function(names = "names element", web_page = "https://nvb.meertens.knaw.nl/naam/is/") {

    names$first_name <- sapply(strsplit(names$Naam, " "), `[`, 1)  # This code should work as a way to extract first names from the ego characteristics dataset
    names$male <- NA
    names$female <- NA

    for (i in 1:nrow(names)) {

        # print(names$first_name[i])

        web_page <- read_html(paste0("https://nvb.meertens.knaw.nl/naam/is/", names$first_name[i]))

        table <- web_page %>%
            rvest::html_elements("body") %>%
            rvest::html_elements("table") %>%
            rvest::html_table()

        if (length(table) == 0) {

            print(length(table))

            names$male[i] <- NA
            names$female[i] <- NA

        } else {

            # print(table) print(table[[1]][[2,3]]) # Check if values for male are coherent and
            # accurate print(table[[1]][[6,3]]) # Check if values for female are coherent and
            # accurate

            names$male[i] <- as.numeric(ifelse(table[[1]][[2, 3]] == "--", 0, table[[1]][[2, 3]]))  # Make sure non-occurences are not registered as '--'
            names$female[i] <- as.numeric(ifelse(table[[1]][[6, 3]] == "--", 0, table[[1]][[6, 3]]))  # Make sure non-occurences are not registered as '--'

        }

    }  # end forloop

    names <- names %>%
        mutate(perc_female = case_when(is.na(female == TRUE) & is.na(male) == TRUE ~ NA, is.na(female) ==
            TRUE ~ 0, is.na(male == TRUE) ~ 1, .default = round((female/(male + female)), 2))) %>%
        select(!c(male, female, first_name))

    return(names)


}  # end function

# ?mutate() ?case_when
df_ego <- gender_scraper.NV(names = df_ego, web_page = "https://nvb.meertens.knaw.nl/naam/is/")
```
```{r}
check <- df_ego %>%
    count(perc_female)

plot(check$perc_female, check$n)
```







