
# Assignments in class: chapter 7
```{r}
#empty environment
rm(list = ls())

#load packages functions
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```
# Packages
```{r}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```
Simulation Logic, steps:

    Sample ego
    Construct possible alternative future networks based on all possible ministeps of ego
    Calculate how sampled ego evaluates these possible networks
    Let the ego pick a network, that is, let agent decide on a tie-change
    GOTO 1 (STOPPING RULE: until you think we have made enough ministeps)

```{r}
#Sample ego
ts_net1
```


```{r}
#plot ts_net1
net1g <- graph_from_adjacency_matrix(ts_net1, mode = "directed")
coords <- layout_(net1g, nicely())  #let us keep the layout
par(mar = c(0.1, 0.1, 0.1, 0.1))
{
    plot.igraph(net1g, layout = coords)
    graphics::box()
}
```


```{r}
#randomly select/sample an agent
set.seed(24553253)
ego <- ts_select(net = ts_net1, steps = 1)  #in rsienatwostep two actors may make a change together but here not
ego
```

```{r}
# a list of all possible networks after all possible tie-changes available to an ego given the current network
options <- ts_alternatives_ministep(net = ts_net1, ego = ego)
# options
plots <- lapply(options, graph_from_adjacency_matrix, mode = "directed")
par(mar = c(0, 0, 0, 0) + 0.1)
par(mfrow = c(1, 2))

fplot <- function(x) {
    plot.igraph(x, layout = coords, margin = 0)
    graphics::box()
}

lapply(plots, fplot)
```

```{r}
#number of ties ego sends to alters
ts_degree(net = options[[1]], ego = ego)
# or for all options
lapply(options, ts_degree, ego = ego)
```
```{r}
#count number of reciprocated ties
lapply(options, ts_recip, ego = ego)
```
```{r}
# ts_degree(options[[option]], ego = ego) --> counts outsending degrees for ego #4
# multiply that by statistic degree: -1  --> ego does not like to send ties
# add: ts_recip(options[[option]], ego = ego) multiplied by reciprocity parameter: 1.5   --> ego likes reciprocated ties
option <- 4 #4 is one of the options that ego #4 has
ts_degree(options[[option]], ego = ego) * -1 + ts_recip(options[[option]], ego = ego) * 1.5
```
```{r}
#evaluation
eval <- ts_eval(net = options[[option]], ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
```
```{r}
#evaluation of all networks
eval <- sapply(options, FUN = ts_eval, ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
print("network with maximum evaluation score:")
which.max(eval)
```
```{r}
# probabilities will sum up to 1: it will either happen or not, total is 1
# translate evaluation score by taking the exponent of them
exp(-0.5)
exp(0.5)
```
```{r}
#normalizing the probabilities

choice <- sample(1:length(eval), size = 1, prob = exp(eval)/sum(exp(eval)))
print("choice:")
choice
#print('network:') options[[choice]]
```

```{r}
#simulate how the network could evolve:
rate <- 2
degree <- -1
recip <- 1.5
ts_sims(nsims = 1, net = ts_net1, startvalues = c(rate, degree, recip), statistics = list(ts_degree,
    ts_recip), p2step = c(1, 0, 0), chain = FALSE)  #not that rate parameter is automatically included. 

#shows one of the possible networks. Do this a trillion times to find the most likely
```
```{r}
#estimation logic
# we do not calculate SE for now.

#ts_targets(net1 = s501, net2 = s502, statistics = list(ts_degree, ts_recip))

# we do not calculate SE for now.
#ans <- ts_estim(net1 = s501, net2 = s502, statistics = list(ts_degree, ts_recip), p2step = c(1, 0, 0),
    #conv = 0.01, phase3 = FALSE, itef3 = 50, verbose = TRUE, startvalues = )

#doesnt work
```
## Interpretation of parameters
```{r}
#degree parameter positive? -> i like degrees! But, we like few ties over many ties.

#it is generally easier to look at ratios of probabilities -- logit

#Rsiena tests at ego level, not macro level --> need to match your methods, relative importance statistics; parameters, simulations
#Keep in mind that changes are dependent on structure of network at t1

```

```{r}
#Rate parameter
par(mfrow = c(1, 3))

dist_5 <- rexp(10000, rate = 5)
hist(dist_5, main = "rate = lambda_i = 5", freq = FALSE, xlab = "waiting times", xlim = c(0, 2), ylim = c(0,
    9))
abline(v = 1/5, col = "red")

dist_10 <- rexp(10000, rate = 10)
hist(dist_10, main = "rate= lambda_j = 10", freq = FALSE, xlab = "waiting times", xlim = c(0, 2), ylim = c(0,
    9))
abline(v = 1/10, col = "red")

dist_15 <- rexp(10000, rate = 15)
hist(dist_10, main = "rate = lambda_k = 15", freq = FALSE, xlab = "waiting times", xlim = c(0, 2), ylim = c(0,
    9))
abline(v = 1/15, col = "red")
```
























