---
title: "Rsiena tutorial with my own data"
author: "by: Frida van Teijlingen"
bibliography: references.bib
---
# Set up
```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, file, "_", datename, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}




packages = c("RSiena", "tidyverse", "stringdist", "stringi")

fpackage.check(packages)
```

# Loading in data
```{r loading in data, echo = F, include = F}
#df_works
df_works <- fload("data/processed/df_works_20241007.rda")

#df_eg
df_ego <- fload("data/processed/df_ego_20241007.rda")

#df_allunis
unidata <- fload("data/processed/unidata_20241007.rda")

```



step 1: define the data
step 2: create the effects structure
step 3: get initial description, look at this file, report something on it (mydata, myeff objects)
        object options, control variables: isolateNet, inPop, outAct
step 4: specify the model (back and forth between 2 & 4) 
step 5: estimate the model, chapter 12 for effects


# Step 1: Define the data
We need to define dependent variables on an ego level or network level. We also need to define independent variables, these can be time constant and time varying.
```{r}
df_uni <- unidata$data
df_network <- unidata$nets

df_demo <- do.call(rbind.data.frame, df_uni$demographics)

```


```{r}
length(df_network)
dim(unidata$nets)#[1]      #1= eerste dimensie, 2=kolom, 3=rij
#2 waves, dus 2 adjacency matrices van elk 234*234
```






# Step 2: Create the effects structure
```{r}

```















